---
title: 'Monster Lab Data Presentation: Wine Preferences'
author: "Julianne Ammirati"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
  github_document: default
  word_document: default
header-includes:
- \usepackage{setspace}
- \doublespacing
---

The code used to generate this presentation is available here: [https://github.com/JulianneA/EDA_w_modeling.git](https://github.com/JulianneA/EDA_w_modeling.git).

```{r setup, include=FALSE}
# loading tidyverse and haven libraries... not sure if this is needed.... 
library(tidyverse)
library(haven)
library(ggplot2)
library(car)
library(stargazer)
```
```{r, include=FALSE}
# read in the comma delimited (CSV) formatted dataset
WINE <- read_csv("White_wines.csv", col_names=TRUE)
```
```{r, include=FALSE}
#Variables (columns) in WINE dataset
names(WINE)
```
## Portugese White Wine Dataset Summary
```{r, include=FALSE}
(summary(WINE))
Qual <-c(round(mean(WINE$quality), digits=2),
round(sd(WINE$quality), digits=2),
median(WINE$quality),
min(WINE$quality),
max(WINE$quality))

NumValues <-c(length(WINE$quality), length(WINE$`fixed acidity`), length(WINE$`volatile acidity`), length(WINE$`citric acid`), length(WINE$`residual sugar`), length(WINE$chlorides),length(WINE$`free sulfur dioxide`), length(WINE$`total sulfur dioxide`), length(WINE$density), length(WINE$pH), length(WINE$sulphates), length(WINE$alcohol), length(WINE$group))

FIXACID <- c(round(mean(WINE$`fixed acidity`), digits=2), round(sd(WINE$`fixed acidity`), digits=2), median(WINE$`fixed acidity`),
min(WINE$`fixed acidity`), max(WINE$`fixed acidity`))

VOLACID <- c(round(mean(WINE$`volatile acidity`), digits=2),
  round(sd(WINE$`volatile acidity`), digits=2),
  median(WINE$`volatile acidity`),
  min(WINE$`volatile acidity`),
  max(WINE$`volatile acidity`))

CITACID <- c(round(mean(WINE$`citric acid`), digits=2),
  round(sd(WINE$`citric acid`), digits=2),
  median(WINE$`citric acid`),
  min(WINE$`citric acid`),
  max(WINE$`citric acid`))
```
```{r, include=FALSE}
RESACID <- c(round(mean(WINE$`residual sugar`), digits=2),
    round(sd(WINE$`residual sugar`), digits=2),
    median(WINE$`residual sugar`),
    min(WINE$`residual sugar`),
    max(WINE$`residual sugar`))

CHLOR <- c(round(mean(WINE$chlorides), digits=2),
    round(sd(WINE$chlorides), digits=2),
    median(WINE$chlorides),
    min(WINE$chlorides),
    max(WINE$chlorides))

FSD <- c(round(mean(WINE$`free sulfur dioxide`), digits=2),
    round(sd(WINE$`free sulfur dioxide`), digits=2),
    median(WINE$`free sulfur dioxide`),
    min(WINE$`free sulfur dioxide`),
    max(WINE$`free sulfur dioxide`))

TSD <- c(round(mean(WINE$`total sulfur dioxide`), digits=2),
    round(sd(WINE$`total sulfur dioxide`), digits=2),
    median(WINE$`total sulfur dioxide`),
    min(WINE$`total sulfur dioxide`),
    max(WINE$`total sulfur dioxide`))

DEN <- c(round(mean(WINE$density), digits=3),
    round(sd(WINE$density), digits=3),
    round(median(WINE$density), digits=3),
    round(min(WINE$density), digits=3),
    round(max(WINE$density), digits=3))

PH <- c(round(mean(WINE$pH), digits=2),
    round(sd(WINE$pH), digits=2),
    median(WINE$pH),
    min(WINE$pH),
    max(WINE$pH))

SULPH <- c(round(mean(WINE$sulphates), digits=2),
    round(sd(WINE$sulphates), digits=2),
    median(WINE$sulphates),
    min(WINE$sulphates),
    max(WINE$sulphates))

ETOH <- c(round(mean(WINE$alcohol), digits=2),
    round(sd(WINE$alcohol), digits=2),
    median(WINE$alcohol),
    min(WINE$alcohol),
    max(WINE$alcohol))

#GRP <- c(round(mean(WINE$group), digits=2),
#    sd(WINE$group),
#    median(WINE$group),
#    min(WINE$group),
#    max(WINE$group))
#Does not seem to recognize column name "group"
```
This dataset contains preference information on `r length(WINE$quality)` different Portugese white varietals covering the following variables: `r names(WINE)`. Overall, preference scores (coded as the _quality_ variable) range from a low score of `r min(WINE$quality)` to a high score of `r max(WINE$quality)` with a mean score of `r round(mean(WINE$quality), digits=2)` and a median score of `r median(WINE$quality)` (SD=`r round(sd(WINE$quality), digits=2)`). The dataset does not contain any missing values. 

**Summary Statistics for all variables are below:**

| | N | Min | Max | Mean | Std Dev | Median |
|-----|------- | ----------|----------| -----------|-----------| ---------|
|Quality |`r NumValues[1]` | `r Qual[4]`|`r Qual[5]`| `r Qual[1]`|`r Qual[2]`|`r Qual[3]`|
| Fixed Acidity |`r NumValues[2]` | `r FIXACID[4]`|`r FIXACID[5]`| `r FIXACID[1]`|`r FIXACID[2]`|`r FIXACID[3]`|
| Volatile Acidity|`r NumValues[3]` | `r VOLACID[4]`|`r VOLACID[5]`| `r VOLACID[1]`|`r VOLACID[2]`|`r VOLACID[3]`|
| Citric Acid |`r NumValues[4]` | `r CITACID[4]`|`r CITACID[5]`| `r CITACID[1]`|`r CITACID[2]`|`r CITACID[3]`|
| Residual Sugar |`r NumValues[5]` | `r RESACID[4]`|`r RESACID[5]`| `r RESACID[1]`|`r RESACID[2]`|`r RESACID[3]`|
| Chlorides |`r NumValues[6]` | `r CHLOR[4]`|`r CHLOR[5]`| `r CHLOR[1]`|`r CHLOR[2]`|`r CHLOR[3]`|
|Free Sulfur Dioxide|`r NumValues[7]` | `r FSD[4]`|`r FSD[5]`| `r FSD[1]`|`r FSD[2]`|`r FSD[3]`|
|Total Sulfur Dioxide|`r NumValues[8]` | `r TSD[4]`|`r TSD[5]`| `r TSD[1]`|`r TSD[2]`|`r TSD[3]`|
| Density |`r NumValues[9]` | `r DEN[4]`|`r DEN[5]`| `r DEN[1]`|`r DEN[2]`|`r DEN[3]`|
| pH |`r NumValues[10]` | `r PH[4]`|`r PH[5]`| `r PH[1]`|`r PH[2]`|`r PH[3]`|
| Sulphates |`r NumValues[11]` | `r SULPH[4]`|`r SULPH[5]`| `r SULPH[1]`|`r SULPH[2]`|`r SULPH[3]`|
| Alcohol |`r NumValues[12]` | `r ETOH[4]`|`r ETOH[5]`| `r ETOH[1]`|`r ETOH[2]`|`r ETOH[3]`|

## Evaluating the Mean Values for Wine Characteristics Based on Quality Rating
Our variable of interest is the wine's quality, and how this variable is impacted by the other characteristics, such as alcohol ocntent or pH. Below is a table of the mean value of each variable when the wines are grouped by quality rating. Each value is rounded to the nearest hundredth, except density measures.
```{r, include=FALSE}
FixedAcidBin <-aggregate(`fixed acidity` ~ quality, WINE, mean)
FixedVolatileBin <-aggregate(`volatile acidity` ~ quality, WINE, mean)
CitricBin <-aggregate(`citric acid` ~ quality, WINE, mean)
SugarBin <-aggregate(`residual sugar` ~ quality, WINE, mean)
ChlorBin <-aggregate(chlorides ~ quality, WINE, mean)
FreeSulf <-aggregate(`free sulfur dioxide` ~ quality, WINE, mean)
TotalSulf <-aggregate(`total sulfur dioxide` ~ quality, WINE, mean)
DensAgg <-aggregate(density ~ quality, WINE, mean)
PHAgg <-aggregate(pH ~ quality, WINE, mean)
SulpAgg <-aggregate(sulphates ~ quality, WINE, mean)
AlcAgg <-aggregate(alcohol ~ quality, WINE, mean)

#DATA <-aggregate(c(WINE$`fixed acidity`, WINE$`volatile acidity`, WINE$`citric acid`,WINE$`residual sugar`, WINE$chlorides, WINE$`free sulfur dioxide`), by=quality, FUN=mean, na.rm=TRUE) #by.y = quality, WINE, mean)

#Hoops to merge into 1 data frame
MERGE1 <-merge(FixedAcidBin, FixedVolatileBin, by.x = "quality")
MERGE2 <-merge(MERGE1, CitricBin, by.x = "quality")
MERGE3 <-merge(MERGE2, SugarBin, by.x = "quality")
MERGE4 <-merge(MERGE3, ChlorBin, by.x = "quality")
MERGE5 <-merge(MERGE4, FreeSulf, by.x = "quality")
MERGE6 <-merge(MERGE5, TotalSulf, by.x = "quality")
MERGE7 <-merge(MERGE6, AlcAgg, by.x = "quality")
MERGE8 <-merge(MERGE7, PHAgg, by.x = "quality")
MERGE9 <-merge(MERGE8, SulpAgg, by.x = "quality")
MeanWine1 <-merge(MERGE9, DensAgg, by.x = "quality")


##CREATE AND MERGE rounded datasets
MERGE10 <-round(MERGE9, 2)
MERGE11 <-round(DensAgg, 4)
MeanWine <-merge(MERGE10, MERGE11, by.x = "quality")
```
``` {r, include=FALSE}
###This is a much easier way to do this, but I don't know how to call on these values for the table below, and the KNITR table is ugly.
FINAL <-WINE %>%
    group_by(quality) %>%
   summarise(MeanFixedAcidity = mean(`fixed acidity`),
              MeanVolatileAcidity = mean(`volatile acidity`),
              MeanCitricAcid = mean(`citric acid`),
              MeanResidualSugar = mean(`residual sugar`),
              MeanChlorides = mean(`chlorides`),
              MeanFreeSulfurDioxide = mean(`free sulfur dioxide`)) %>%
    knitr::kable(col.names=c("Quality", "Mean Fixed Acidity", "Mean Volatile Acidity", 
                    "Mean Citric Acid","Mean Residual Sugar", "Mean Chlorides", 
                    "Mean Free Sulfur Dioxide"),
                digits = 2,
                 caption="Mean Values for Wine Characteristics By Quality Rating")
```
```{r, include=FALSE}
#knitr::kable(MERGE) ## It looks better when typed out in RMD
```
|Quality | `r MeanWine[1,1]`| `r MeanWine[2,1]` |`r MeanWine[3,1]`| `r MeanWine[4,1]`|`r MeanWine[5,1]`|`r MeanWine[6,1]`|`r MeanWine[7,1]`|
|-----|------- | ----------|----------| -----------|-----------| ---------|
| Fixed Acidity | `r MeanWine[1,2]`| `r MeanWine[2,2]` |`r MeanWine[3,2]`| `r MeanWine[4,2]`|`r MeanWine[5,2]`|`r MeanWine[6,2]`|`r MeanWine[7,2]`|
| Volatile Acidity| `r MeanWine[1,3]`| `r MeanWine[2,3]` |`r MeanWine[3,3]`| `r MeanWine[4,3]`|`r MeanWine[5,3]`|`r MeanWine[6,3]`|`r MeanWine[7,3]`|
| Citric Acid | `r MeanWine[1,4]`| `r MeanWine[2,4]` |`r MeanWine[3,4]`| `r MeanWine[4,4]`|`r MeanWine[5,4]`|`r MeanWine[6,4]`|`r MeanWine[7,4]`|
| Residual Sugar | `r MeanWine[1,5]`| `r MeanWine[2,5]` |`r MeanWine[3,5]`| `r MeanWine[4,5]`|`r MeanWine[5,5]`|`r MeanWine[6,5]`|`r MeanWine[7,5]`|
| Chlorides | `r MeanWine[1,6]`| `r MeanWine[2,6]` |`r MeanWine[3,6]`| `r MeanWine[4,6]`|`r MeanWine[5,6]`|`r MeanWine[6,6]`|`r MeanWine[7,6]`|
|Free Sulfur Dioxide| `r MeanWine[1,7]`| `r MeanWine[2,7]` |`r MeanWine[3,7]`| `r MeanWine[4,7]`|`r MeanWine[5,7]`|`r MeanWine[6,7]`|`r MeanWine[7,7]`|
|Total Sulfur Dioxide| `r MeanWine[1,8]`| `r MeanWine[2,8]` |`r MeanWine[3,8]`| `r MeanWine[4,8]`|`r MeanWine[5,8]`|`r MeanWine[6,8]`|`r MeanWine[7,8]`|
| Density | `r MeanWine[1,9]`| `r MeanWine[2,9]` |`r MeanWine[3,9]`| `r MeanWine[4,9]`|`r MeanWine[5,9]`|`r MeanWine[6,9]`|`r MeanWine[7,9]`|
| pH | `r MeanWine[1,10]`| `r MeanWine[2,10]` |`r MeanWine[3,10]`| `r MeanWine[4,10]`|`r MeanWine[5,10]`|`r MeanWine[6,10]`|`r MeanWine[7,10]`|
| Sulphates | `r MeanWine[1,11]`| `r MeanWine[2,11]` |`r MeanWine[3,11]`| `r MeanWine[4,11]`|`r MeanWine[5,11]`|`r MeanWine[6,11]`|`r MeanWine[7,11]`|
| Alcohol | `r MeanWine[1,12]`| `r MeanWine[2,12]` |`r MeanWine[3,12]`| `r MeanWine[4,12]`|`r MeanWine[5,12]`|`r MeanWine[6,12]`|`r MeanWine[7,12]`|

## Means of All Independent Variables by Quality Value
These graphs are histograms of all data points for the given characteristic. The hot-pink overlay tracks the mean value at each quality level for that characteristic, which is purely exploratory and not meant to display significant patterns in the data.
```{r, include=FALSE}
#rename column headers to indicate that these are means
names(MeanWine1) <- c("Quality", "Mean Fixed Acidity", "Mean Volatile Acidity", 
                  "Mean Citric Acid","Mean Residual Sugar", "Mean Chlorides", 
                  "Mean Free Sulfur Dioxide", "Mean Total Sulfur Dioxide", "Mean Alcohol", 
                  "Mean pH", "Mean Sulphates", "Mean Density")
```
```{r, echo=FALSE}
ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Fixed Acidity`))+
    geom_line(stat="identity", color='hotpink', size=2)+
    #geom_boxplot(data=WINE,aes(x=quality, y=`fixed acidity`)) +
    geom_jitter(data=WINE, aes(x=quality, y=`fixed acidity`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Volatile Acidity`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
  geom_jitter(data=WINE, aes(x=quality, y=`volatile acidity`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Citric Acid`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
  geom_jitter(data=WINE, aes(x=quality, y=`citric acid`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Residual Sugar`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
geom_jitter(data=WINE, aes(x=quality, y=`residual sugar`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Chlorides`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
geom_jitter(data=WINE, aes(x=quality, y=`chlorides`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Free Sulfur Dioxide`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
geom_jitter(data=WINE, aes(x=quality, y=`free sulfur dioxide`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Total Sulfur Dioxide`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
geom_jitter(data=WINE, aes(x=quality, y=`total sulfur dioxide`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Density`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
geom_jitter(data=WINE, aes(x=quality, y=`density`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean pH`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
geom_jitter(data=WINE, aes(x=quality, y=`pH`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Sulphates`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
geom_jitter(data=WINE, aes(x=quality, y=`sulphates`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)

ggplot(data=MeanWine1, aes(x=Quality, y=`Mean Alcohol`)) +
    geom_line(stat="identity", color='hotpink', size=2)+
geom_jitter(data=WINE, aes(x=quality, y=`alcohol`), position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
```

```{r, include=FALSE}
#Opted for line graphs (above) because these are difficult to discern because of # of data points...
WINE %>%
  ggplot(aes(x=`fixed acidity`, y=quality)) +
    geom_point()
WINE %>%
  ggplot(aes(x=`volatile acidity`, y=quality)) +
    geom_point()
WINE %>%
  ggplot(aes(x=`citric acid`, y=quality)) +
    geom_point()
WINE %>%
  ggplot(aes(x=`residual sugar`, y=quality)) +
    geom_point()
WINE %>%
  ggplot(aes(x=chlorides, y=quality)) +
    geom_point()
WINE %>%
  ggplot(aes(x=`free sulfur dioxide`, y=quality)) +
    geom_point()
```
```{r, echo=FALSE, include=FALSE}
#this is spliting the data into test and train groups.
set.seed(20170212) #This can be anything
WINE$group <- runif(length(WINE$quality), min = 0, max = 1)
summary(WINE)
Wine.train <- subset(WINE, group <= 0.90) #Subset(Dataset, variable less than or equal to .90)
#This splits the data 90/10 and we can do it because there is such a large dataset
Wine.test <- subset(WINE, group > 0.90)
```
```{r, include=FALSE}
plot(quality ~ `fixed acidity`, data = Wine.test) #y ~ x (outcome(y) depends on predictors(x))
plot(quality ~ `volatile acidity`, data = Wine.test)  
plot(quality ~ `citric acid`, data = Wine.test) 
plot(quality ~ `residual sugar`, data = Wine.test)  
plot(quality ~ chlorides, data = Wine.test) 
plot(quality ~ `free sulfur dioxide`, data = Wine.test) 
plot(quality ~ `total sulfur dioxide`, data = Wine.test) 
plot(quality ~ density, data = Wine.test) 
plot(quality ~ pH, data = Wine.test) 
plot(quality ~ sulphates, data = Wine.test) 
plot(quality ~ alcohol, data = Wine.test) 

``` 

##Univariate Modeling
Based on the histographs and line graphs above, alcohol content appears to be the strongest candidate for a significant linear relationship with the quality measure.

Exploratory simple linear modeling was conducted to complete a preliminary examination of the independent variable by quality relationship. While many of the linear models were found to be significant, this is likely the result of the sample size rather than true linear relationship between variables.
```{r, include=FALSE}
# linear regression of effect on fixed acidity on quality
regfixacid <- lm(quality ~ `fixed acidity`, data = Wine.train) #lm is linear model function
#This will give us the linear model, but by assigning it to reg1, we can save results in an object
summary(regfixacid)
#repeating with other variables
regvolacid <- lm(quality ~ `volatile acidity`, data = Wine.train)
regcitacid <- lm(quality ~ `citric acid`, data = Wine.train)
regressugar <- lm(quality ~ `residual sugar`, data = Wine.train)
regchlor <- lm(quality ~ chlorides, data = Wine.train)
regfsd <- lm(quality ~ `free sulfur dioxide`, data = Wine.train)
regtsd <- lm(quality ~ `total sulfur dioxide`, data = Wine.train)
regden <- lm(quality ~ density, data = Wine.train)
regpH <- lm(quality ~ pH, data = Wine.train)
regsulph <- lm(quality ~ sulphates, data = Wine.train)
regetoh <- lm(quality ~ alcohol, data = Wine.train)

summary(regvolacid) #SIGNIFICANT
summary(regcitacid)
summary(regressugar)#SIGNIFICANT
summary(regchlor)#SIGNIFICANT
summary(regfsd)
summary(regtsd)#SIGNIFICANT
summary(regden)#SIGNIFICANT
summary(regpH)
summary(regsulph)
summary(regetoh)#SIGNIFICANT
```

From this univariate analysis, six variables were found to have a significant linear relationship (p<0.001) with wine quality. Volatile acidity, residual sugar, chlorides, total sulfur dioxide, density, and alcohol content.

* For each increase of 1 unit in **Fixed Acidity**, the quality score of the wine decreases by 0.12

* For each increase of 1 unit in **Volatile Acidity**, the quality score of the wine decreases by 1.75

* For each increase of 10 units in **Residual Sugar**, the quality score of the wine decreases by 0.16

* For each increase of 1/10 unit in **Chlorides**, the quality score of the wine decreases by 8.31

* For each increase of 100 unit in **Total sulfur dioxide**, the quality score of the wine decreases by 0.40

* For each increase of 1/100 unit in **Density**, the quality score of the wine decreases by 0.90

* For each increase of 1 unit in **pH**, the quality score of the wine increases by 0.59

* For each increase of 1 unit in **Sulphates**, the quality score of the wine increases by 0.40

* For each increase of 1 unit in **Alcohol content**, the quality score of the wine increases by 0.31

```{r, include=FALSE}
Wine.train$pred.qual <- predict(regetoh, newdata = Wine.train) #predict function will use the linear model object
Wine.test$pred.qual <- predict(regetoh, newdata=Wine.test)

# calculate residuals in the training and testing dataset
Wine.train$resid <- Wine.train$quality - Wine.train$pred.qual
Wine.test$resid <- Wine.test$quality - Wine.test$pred.qual
```

```{r, include=FALSE}
library(car)
#get the residual plots
residualPlots(regetoh)
#gives you p-values that aren't statistically significant
```

```{r, include=FALSE}
#plot test dataset
plot(quality ~ alcohol, data = Wine.test) 
```
```{r, include=FALSE}
#OR CAN JUST RUN IT LIKE THIS
plot(quality~alcohol, data = Wine.test)
abline (0,0)
```

##Multivariate Modeling
####Scatterplot Matrix of variables in Wine dataset:

```{r, echo=FALSE}
##I HAVE NO IDEA WHAT I AM DOING HERE.....
#produce a scatterplot matrix 
scatterplotMatrix(~ quality + alcohol + density + `residual sugar` +`volatile acidity`, span =0.7, data = WINE)
```
```{r, include=FALSE}
#plotting variable by variable scatterplots
library(car)
plot(WINE$density, WINE$alcohol)
plot(WINE$alcohol, WINE$`residual sugar`)
plot(WINE$alcohol, WINE$`volatile acidity`)

#WINE$DensityRevised  <- factor(WINE$density,
                            #levels = c(1,2,3),
                            #labels = c("No-use","Long-term","Short-term"))

#text(wine$V4, wine$V5, wine$V1, cex=0.7, pos=4, col="red")
```

It is clear from the initial plots that outliers are an issue in this dataset.

```{r, include=FALSE}
##AGAIN, I HAVE NO IDEA WHAT I AM DOING HERE. I am unsure what transformations are worthwhile on this data.
#produce a scatterplot matrix 
scatterplotMatrix(~ quality + alcohol + log2(density) + log2(chlorides), span =0.7, data = WINE)
```
```{r, include=FALSE}
# let the multivariate regressions begin!
##Running with everything... mostly just want to check alcohol...
wine.bigmod <- lm(quality ~ `fixed acidity`+`volatile acidity`+ `citric acid`+`residual sugar`+ chlorides+ `free sulfur dioxide`+`total sulfur dioxide`+ density+ pH+ sulphates+ alcohol, data= Wine.train)

#and see what we have
summary(wine.bigmod)
###Literally, EVERYTHING WAS SIGNIFICANT. Ignore this adventure and try again.
#Fstatistic gives overall F for H0=B1=B2=0

wine.LessBigMod <- lm(quality ~ `fixed acidity`+`volatile acidity`+`residual sugar`+ `free sulfur dioxide`+ density+ pH+ sulphates+ alcohol, data= Wine.train)

#and see what we have
summary(wine.LessBigMod)
#Fstatistic gives overall F for H0=B1=B2=0

```
```{r, include=FALSE}
wine.ILikeThisMod <- lm(quality ~ alcohol+density, data= Wine.train)

#and see what we have
summary(wine.ILikeThisMod)
##Did not like as much as I had hoped
```

It initially seems as though alcohol content has some significant relationship to the quality of the wine. MLR models were generated to evauate which characteristicis improved the variation in quality accounted for by the model (optimized for adjusted R-squared value). Models were iterated on to continue to improve R-squared, with additional variables being gradually added to the strongest initial model: quality ~ alcohol.

* At most, these models were only able to account for approximately 26% of the variation in quality among the various wines. 

####Optimized model for this dataset
```{r, include=FALSE}
#I do not know enough about statistics to actually know what I am doing. I am sure there is an easier way, but I completed an exhaustive model comparison, and the best I could get was an adjusted R-value of 26%
#wine.mod2 <- lm(quality ~ alcohol+ fixed.acidity + volatile.acidity , data=WINE)
#wine.mod3 <- lm(quality ~ alcohol+free.sulfur.dioxide+total.sulfur.dioxide, data= WINE)
#wine.mod4 <- lm(quality ~ alcohol+density+ pH, data= WINE)
wine.mod5 <- lm(quality ~ alcohol+`fixed acidity`, data= Wine.train)
wine.mod6 <- lm(quality ~ alcohol+`volatile acidity`, data= Wine.train)
wine.mod7 <- lm(quality ~ alcohol+`residual sugar`, data= Wine.train)
wine.mod8 <- lm(quality ~ alcohol+`free sulfur dioxide`, data= Wine.train)
wine.mod9 <- lm(quality ~ alcohol+`total sulfur dioxide`, data= Wine.train)
#wine.mod5 <- lm(quality ~ alcohol+ fixed.acidity , data= WINE)
#wine.mod6 <- lm(quality ~ alcohol+ volatile.acidity , data= WINE)
#wine.mod7 <- lm(quality ~ alcohol+ residual.sugar , data= WINE)
#wine.mod8 <- lm(quality ~ alcohol+free.sulfur.dioxide, data= WINE)
#wine.mod9 <- lm(quality ~ alcohol+total.sulfur.dioxide, data= WINE)
wine.mod10 <- lm(quality ~ alcohol+density, data= Wine.train)
wine.mod11 <- lm(quality ~ alcohol+pH, data= Wine.train)
wine.mod12 <- lm(quality ~ alcohol+sulphates, data= Wine.train)

summary(wine.mod5)#Adjusted R-squared is 0.1955
summary(wine.mod6)#Adjusted R-squared is 0.243
summary(wine.mod7)#Adjusted R-squared is 0.2049
summary(wine.mod8)#Adjusted R-squared is 0.2057
summary(wine.mod9)#Adjusted R-squared is 0.1917
summary(wine.mod10)#Adjusted R-squared is 0.1947
summary(wine.mod11)#Adjusted R-squared is 0.1939
summary(wine.mod12)#Adjusted R-squared is 0.1948

#VOLATILE ACIDITY IMPROVED MODEL ThE MOST AFTER ALCOHOL.
```
```{r, include=FALSE}

#Looking to improve model beyond 0.243
wine.mod13 <- lm(quality ~ alcohol+`volatile acidity`+`fixed acidity`, data= Wine.train) 
wine.mod14 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`, data= Wine.train) 
wine.mod15 <- lm(quality ~ alcohol+`volatile acidity`+`free sulfur dioxide`, data= Wine.train) 
wine.mod16 <- lm(quality ~ alcohol+`volatile acidity`+`total sulfur dioxide`, data= Wine.train) 
wine.mod17 <- lm(quality ~ alcohol+`volatile acidity`+density, data= Wine.train) 
#wine.mod13 <- lm(quality ~ alcohol+density+fixed.acidity, data= WINE) #0.1982
#wine.mod14 <- lm(quality ~ alcohol+density+volatile.acidity, data= WINE) #0.2464
#wine.mod15 <- lm(quality ~ alcohol+density+residual.sugar, data= WINE) #0.2097
#wine.mod16 <- lm(quality ~ alcohol+density+free.sulfur.dioxide, data= WINE) #0.2049
#wine.mod17 <- lm(quality ~ alcohol+density+total.sulfur.dioxide, data= WINE) #0.192
wine.mod18 <- lm(quality ~ alcohol+`volatile acidity`+pH, data= Wine.train) #0.1942
wine.mod19 <- lm(quality ~ alcohol+`volatile acidity`+sulphates, data= Wine.train) #0.1951
summary(wine.mod13) #Adjusted R-squared is 0.2472
summary(wine.mod14) #Adjusted R-squared is 0.2625
summary(wine.mod15) #Adjusted R-squared is 0.253 
summary(wine.mod16) #Adjusted R-squared is 0.245
summary(wine.mod17) #Adjusted R-squared is 0.2501
summary(wine.mod18) #Adjusted R-squared is 0.2443
summary(wine.mod19) #Adjusted R-squared is 0.2453

#Residual sugar improved prediction the most, to 0.2625
```
```{r, include=FALSE}
#Residual sugar improved prediction the most, to 0.2625. Looking for new model that improves beyond 0.2625
wine.mod20 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`+`fixed acidity`, data= Wine.train)
wine.mod21 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`+`citric acid`, data= Wine.train)
wine.mod22 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`+chlorides, data= Wine.train)
wine.mod23 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`+`free sulfur dioxide`, data= Wine.train) 
wine.mod24 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`+`total sulfur dioxide`, data= Wine.train) 
wine.mod25 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`+density, data= Wine.train)
wine.mod26 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`+pH, data= Wine.train)
wine.mod27 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`+sulphates, data= Wine.train) 
#wine.mod20 <- lm(quality ~ alcohol+density+volatile.acidity+fixed.acidity, data= WINE) #0.2546
#wine.mod21 <- lm(quality ~ alcohol+density+volatile.acidity+residual.sugar, data= WINE) #0.2633
#wine.mod22 <- lm(quality ~ alcohol+density+volatile.acidity+free.sulfur.dioxide, data= WINE) #0.2543
#wine.mod23 <- lm(quality ~ alcohol+density+volatile.acidity+pH, data= WINE) #0.2477
#wine.mod24 <- lm(quality ~ alcohol+density+volatile.acidity+sulphates, data= WINE) #0.2484
summary(wine.mod20) #Adjusted R-squared is 0.2676
summary(wine.mod21) #Adjusted R-squared is 0.2627
summary(wine.mod22) #Adjusted R-squared is 0.2627
summary(wine.mod23) #Adjusted R-squared is 0.2673
summary(wine.mod24) #Adjusted R-squared is 0.2625
summary(wine.mod25) #Adjusted R-squared is 0.2678
summary(wine.mod26) #Adjusted R-squared is 0.2661
summary(wine.mod27) #Adjusted R-squared is 0.2654

#Residual sugar improved prediction to 0.2625. Density improved the model further to 0.2678

```
```{r, include=FALSE}
#Model using alcohol, volatile acidity, residual sugar, and density
wine.mod25 <- lm(quality ~ alcohol+`volatile acidity`+`residual sugar`+density, data= Wine.train)#0.2678
summary(wine.mod25)
```
```{r, include=FALSE}
library(stargazer)
stargazer(wine.mod21,wine.mod14,title="Comparison of 2 Regression outputs",type="pdf",align=TRUE)
```
The final model included alcohol, density, volatile acidity, and residual sugar. This model had an adjusted R-square of 26.78% with the training dataset.

* For each increase of 1 unit in **alcohol content**, the quality score of the wine increases by 0.29, as long as density, volatile acidity and residual sugar were kept constant (p<0.001)

* For each increase of 1/100 unit in **density**, the quality score of the wine decreases by 0.70, as long as alcohol content, volatile acidity and residual sugar were kept constant (p<0.001)

* For each increase of 1 unit in **volatile acidity**, the quality score of the wine decreases by 2.08, as long as alcohol content, density and residual sugar were kept constant (p<0.001)

* For each increase of 10 units in **residual sugar**, the quality score of the wine increases by 0.53, as long as alcohol content, density and volatile acidity were kept constant (p<0.001)

```{r, include=FALSE}
#Used model 21 from above!
wine.finmodel <- lm(quality ~ alcohol+density+`volatile acidity`+`residual sugar`, data= Wine.train)
#wine.finmodel <- lm(quality ~ alcohol+density+volatile.acidity +residual.sugar, data= Wine.train)
summary(wine.finmodel)
```

```{r, include=FALSE}
Wine.train$pred <- predict(wine.mod25, 
                      newdata=Wine.train)#, 
                      #type="response")
Wine.test$pred <- predict(wine.mod25, 
                      newdata=Wine.test)#, 
                      #type="response")
```
```{r, include=FALSE}
# get predicted values in the training and testing dataset
Wine.train$pred.qual <- predict(wine.mod25, newdata = Wine.train) #predict function will use the linear model object
Wine.test$pred.qual <- predict(wine.mod25, newdata=Wine.test)

# calculate residuals in the training and testing dataset
Wine.train$resid <- Wine.train$quality - Wine.train$pred.qual
Wine.test$resid <- Wine.test$quality - Wine.test$pred.qual
```
```{r, include=FALSE}
library(car)
#get the residual plots
residualPlots(wine.mod25)
#gives you p-values that aren't statistically significant
```
```{r, include=FALSE}
#plot test dataset
plot(quality ~ alcohol+`volatile acidity`+`residual sugar`+density, data= Wine.test)
```

```{r, eval=FALSE, include=FALSE}
#This is from regression lecture. I have no idea what its doing.
#set up the 0 line
x0 <- 1 #set start of line
parent <- seq(x0, 1001, 1) # create a sequence from x0 (64) to 73, in 0.01 increments
resid <- rep(0, length(parent)) #draws line (resid) at a y of 0 for the length of "parent"

#plot residual versus X without parameters... its crappy looking
plot(resid~parent, data = Wine.test)
#overlay the zero line
par(new=TRUE) 
plot(resid~parent)#, pch = 16, cex = .1, axes = FALSE, labels = FALSE) #This plots the actual overlay

#set up the 0 line
x0 <- 64
parent <- seq(x0, 73, 0.01) 
resid <- rep(0, length(parent))

#plot residual versus X
plot(resid~parent, data = Galton.test)
#overlay the zero line
par(new=TRUE) #Allows for overlay
plot(resid~quality, pch = 16, cex = .1, axes = FALSE, labels = FALSE) #These parameters set up width of zero line, removes additional labels on y-axis
#This plot is to visualize how the residuals plot around zero- check for homosceda...
```
##Conclusion
Based on the final multivariare regression model using these variables, it is clear that we must reject the null hypothesis. The most comprehensive model that was pulled from this data included four variables in the regression equation: residual sugar, volatile acidity, alcohol content, and density. These four variables interact to predict quality via the following equation:
\begin{equation}
  y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_3
\end{equation}
\begin{equation}
  quality = \beta_0 + [.29*(alcohol)] + [-70.33*(density)] + [-2.08*(volatile acidity)] + [0.05*(residual_sugar)]
\end{equation}

These findings suggest that as alcohol content increases, the quality of the wine is likely to improve as well. Density and volatile acid have inverse relationships with qualtiy, so when these variables increase, the quality will most likely decrease. Finally, residual sugar, like alcohol, has a positive linear relationship with wine quality, so sugar amount and quality are more likely to change value in the same direction. 

It is important to note that models using this data can predict 26% of the variation, at best. In order to improve predictive value of models, it is necessary to clean the data and determine whether certain variables require transformations before performing regression modeling.